To start with, we should tell what machine learning is. 
One of the definitions described it as the field of study that gives a computer the ability to ,,learn to solve problems without being explicitly programmed" \cite{koza}. 
Machine learning problems are categorized into two groups: supervised learning and unsupervised learning. 
The difference between these two is that in supervised learning we already have a data set and correct output. 
In unsupervised learning we just want to find relationships in data and based on it divide it into clusters.
%opisać szczegółowo nadzorowane i nie i machine learning
\par
The classification problem is a supervised learning problem, where we want to predict to which classes a new query fits best. 
There are many well-known algorithms to solve this problem but when we think about a lot of classes, e.g. millions, these algoritms often need a lot of time to find an exact solution. 
Applications of classification can be found in recomendation systems and on-line advertising.
\par
We focused on a problem called Maximum Inner Product Search and it is stated as follows: 
\[ \argmax_{x \in S} q^T x \]
It is a classification problem where we want to classify a huge number of items $q$  to a huge number of classes $x$ belonging to set $S$. 
%mips szerzej
Often there is no need to have an exact result. 
Some vectors having high inner product with query may be good enough. 
To speed up we try to approximate result. 
It is important to find a balance between speedup and precision. 
In this work we will focus on three algorithms which can solve this problem.
